---
title: "Calculus Workshop"
subtitle: "Integration Problem Set"
author: "Nathan Grimes"
date: "2023-09-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE,message=FALSE)

library(tidyverse)
```

*Answer the following questions to the best of your ability. Feel free to work with anyone in the cohort, though I would encourage attempting on your own first to make sure you fully understand the concepts.*

# Logarithms

## 1) Solve for x

$ln(x+2)=12$

\textcolor{blue}{Solution}

\begin{align}
x+2=e^{12}\\
x=e^{12}-2\\
x=162,752.79
\end{align}

$e^{3x+1}=16$

\textcolor{blue}{Solution}

\begin{align}
3x+1=\ln(16)\\
x=\frac{\ln(16)-1}{3} \\
x=1.26
\end{align}

$6e^{4x}=41e^{2x}$

\textcolor{blue}{Solution}

\begin{align}
\frac{e^{4x}}{e^{2x}=\frac{41}{6}}\\
e^{4x-2x}=\frac{41}{6} \\
e^{2x}=\frac{41}{6}\\
2x=\ln(41)-\ln(6)\\
x=\frac{\ln(41)-\ln(6)}{2}
\end{align}


## 2) Find the derivative:

$$
G(a)=\frac{2\ln(4a-2)}{e^{3a}}
$$

\textcolor{blue}{Solution: Break it down into parts. First we need the quotient rule. Then use chain rules for the f(x) and g(x) that you identify. Put it all together}

\begin{align}
\frac{dG}{da}=\frac{\frac{8e^{3a}}{4a-2}-3e^{3a}\ln(4a-2)}{[e^{3a}]^2}
\end{align}



# Reimann Sum

1) Approximate the area under the curve $f(x)=3x^2-6x+10$ on the interval $[6,12]$ with 6 uniform rectangles. 

Does the position of the rectangle make a difference? Describe in words,show mathematically, or draw on the graph how evaluating the rectangle in different ways might lead to slightly different approximations.

```{r}
x=seq(4,14,by=0.1)

y=3*x^3-6*x+10

ggplot()+
  geom_line(aes(x=x,y=y),color="black",linewidth=2)+
  theme_classic()+
  scale_x_continuous(breaks=seq(4,14,by=1))+
  labs(x="",y="")+
  theme(text=element_text(size=19))
  
```

\textcolor{blue}{The position of the rectangles will slightly affect the approximation, but not much. We could have the squares evaluate at the midpoint of $\delta x$ instead of at the end. Or we could evaluate to the right.}

\textcolor{blue}{Evaluate the function at all values in between the interval. Then add them up.}

```{r, echo=TRUE}
x=seq(6,12)

f=3*x^3-6*x+10

sum(f)


```


# Intergrals


## 1)
The marginal benefit of abatement (e.g. reducing) for carbon is given by:

$$
MB=31-2Q
$$

However there is also a marginal cost for carbon abatement given by:

$$
MC=6+3Q
$$

Find the total net benefit of carbon abatement to society at equilibrium. (Hint: To get equilibrium and the bounds of the integral, first set marginal benefit equal to marginal cost and solve for $Q^*$. Then your integral bounds should be from 0 to $Q^*$)

\textcolor{blue}{Solution: First find the equilibrium by setting the MB and MC equations equal to each other and solve for Q}

\begin{align}
31-2Q=6+3Q\\
25=5Q \\
5=Q^*
\end{align}

\textcolor{blue}{Now we can find the total benefit and cost of abatement by taking the definite integral of each curve from 0 to 5}
\textcolor{blue}{Marginal Benefit}

\begin{align}
\int^5_030-2Qdq \\
30Q-Q^2|^5_0 \\
30(5)-(5)^2=125
\end{align}

\textcolor{blue}{Marginal Cost}

\begin{align}
\int^5_06+3Q dq\\
6Q+\frac{3x^2}{2}|^5_0\\
6(5)+\frac{3(5)^2}{2}=67.5
\end{align}

\textcolor{blue}{Now we subtract the difference to get the total net benefits to society (in ESM 204 this will be called net welfare). $125-67.5$=57.5}

## 2)

Take the Integrals

\begin{align}
\text{A) }y=\frac{3}{x^2}, y(1)=5&  &\text{B)  }g(t)=3t^5-2t^3+16t-7  &  &\text{C) } \int^4_2\frac{1}{2}x
\end{align}

\textcolor{blue}{Solution:}

A) 
\begin{align}
\int\frac{3}{x^2} dx\\
-3x^{-3}+C\\
-3(1)^{-3}+C=5\\
C=8\\
x^{-3}+8
\end{align}

B)
\begin{align}
\int\ g(t)=3t^5-2t^3+16t-7 dt\\
\frac{t^6}{2}-\frac{t^4}{2}+8t^2-7t+C\\
\end{align}

C)
\begin{align}
\int^4_2\frac{1}{2}x dx\\
\frac{1}{4}x^2|^4_2\\
\frac{1}{4}(4)^2-\frac{1}{4}(2)^2=3
\end{align}

## 3)

A model for the rate of change in ozone concentrations over time between 1962-1984 is given by $\frac{dC}{dt}=2t+20$. Where $C$ is the ozone concentration (ppm) and t is the elapsed time in years. Given that in 1964 the ozone concentration was 30 ppm, what was the ozone concentration in 1982?

\begin{align}
\frac{dC}{dt}&=2t+20\\
dC&=2t+20 dt \\
\int dC&=\int 2t+20 dt \\
C&=t^2+20t+D\\
30&=2^2+20(2)+D \\
D&=-14
\end{align}

Now that we know D we can solve for the concentration in 1982

$$
20^2+20*(20)-14=\text{ 786ppm}
$$

