---
title: "Calculus Workshop"
subtitle: "Integration Problem Set"
author: "Nathan Grimes"
date: "2023-09-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE,message=FALSE)

library(tidyverse)
```

*Answer the following questions to the best of your ability. Feel free to work with anyone in the cohort, though I would encourage attempting on your own first to make sure you fully understand the concepts.*

# Logarithms

## 1) Solve for x

$ln(x+2)=12$

$e^{3x+1}=16$

$6e^{4x}=41e^{2x}$

## 2) Find the derivative:

$$
G(a)=\frac{2\ln(4a-2)}{e^{3a}}
$$


# Reimann Sum

1) Approximate the area under the curve $f(x)=3x^2-6x+10$ on the interval $[6,12]$ with 6 uniform rectangles. 

Does the position of the rectangle make a difference? Desribe in words,show mathmatically, or draw on the graph how evaluating the rectangle in different ways might lead to slighly different approximations.

```{r}
x=seq(4,14,by=0.1)

y=3*x^3-6*x+10

ggplot()+
  geom_line(aes(x=x,y=y),color="black",linewidth=2)+
  theme_classic()+
  scale_x_continuous(breaks=seq(4,14,by=1))+
  labs(x="",y="")+
  theme(text=element_text(size=19))
  
```

# Intergrals


## 1)
The marginal benefit of abatement (e.g. reducing) for carbon is given by:

$$
MB=31-2Q
$$

However there is also a marginal cost for carbon abatement given by:

$$
MC=6+3Q
$$

Find the total net benefit of carbon abatement to society at equilibrium. (Hint: To get equilibrium and the bounds of the integral, first set marginal benefit equal to marginal cost and solve for $Q^*$. Then your integral bounds should be from 0 to $Q^*$)

## 2)

Take the Integrals

\begin{align}
\text{A) }y=\frac{3}{x^2}, y(0)=5&  &\text{B)  }g(t)=3t^5-2t^3+16t-7  &  &\text{C) } \int^4_2\frac{1}{2}x
\end{align}


## 3)

A model for the rate of change in ozone concentrations over time between 1962-1984 is given by $\frac{dC}{dt}=2t+20$. Where $C$ is the ozone concentration (ppm) and t is the elapsed time in years. Given that in 1964 the ozone concentration was 30 ppm, what was the ozone concentration in 1982?
